Invariant Representations for Noisy Speech Recognition

### 2016-10-22 paper.tex
Changes in model description

% Title.
% ------
\title{INVARIANT REPRESENTATIONS FOR NOISY SPEECH RECOGNITION}
%
% Single address.
% ---------------
%\name{Dmitriy Serdyuk $^{\dagger}$\thanks{Dmitriy Serdyuk performed the work
%  while at IBM}, Kartik Audhkhasi$^{\star}$, Phil\'emon Brakel$^{\dagger}$, Bhuvana Ramabhadran$^{\star}$, Yoshua Bengio$^{\dagger}$\thanks{Yoshua Bengio is a CIFAR Fellow.}}
%\address{$^{\dagger}$ Universit\'e de Montr\'eal\\
%         $^{\star}$ IBM}
\author{
  Dmitriy Serdyuk\thanks{Dmitriy Serdyuk performed the work 
    during an internship at IBM} \\
  MILA\\
  Universit\'e de Montr\'eal\\
  Montr\'eal, QC H3T 1J4 \\
  \texttt{serdyuk@iro.umontreal.ca} \\
  \And
  Kartik Audhkhasi \\
  IBM \\
  Yorktown Heights, NY 10598\\
  \texttt{kaudhkha@us.ibm.com} \\
  \And
  Phil\'emon Brakel \\
  MILA\\
  Universit\'e de Montr\'eal\\
  Montr\'eal, QC H3T 1J4 \\
  \texttt{pbpop3@gmail.com} \\
  \And
  Bhuvana Ramabhadran\\
  IBM \\
  Yorktown Heights, NY 10598\\
  \texttt{bhuvana@us.ibm.com} \\
  \And
  Yoshua Bengio\\
  MILA, CIFAR Fellow\\
  Universit\'e de Montr\'eal\\
  Montr\'eal, QC H3T 1J4 \\
  \texttt{findme@inter.net} \\
}

@@

    %One of the properties of GAN networks that there is no straightforward way 
    %to evaluate the quality of image generation process since the GAN doesn't 
    %output the log-likelihood of its samples. The paper estimated the likelihood
    %using the Parzen windows estimator.

@@

    The network is trained with three goals: the hidden representation $h$ should
    be helpful for the target classifier, but harmful for the domain classifier,
    the domain classifier should have a good classification accuracy. More 
    formally, we define the objective as
    \begin{equation}
        L = L(\hat{y}, y; \theta_R, \theta_C) + 
        \alpha L(\hat{d}, d; \theta_D) -
        \beta L(\hat{d}, d; \theta_R),
    \end{equation}
    where $y$ is the ground truth class, $d$ is the domain label, corresponding
    hatted variables are the network predictions, $\theta$ are the network 
    parameters, and $R, C, D$ are the subsets of parameters in the representation,
    classifier and domain adaptation networks respectively. The hyper-parameters
    $\alpha$ and $\beta$ denote the influence of every part of the cost.

\subsection{Invariant represenations for speech recognition}
\label{sec:invariant-speech}

\subsection{Adversarial Cost Types}

%TODO: explain -log(\hat{y}) and log(1 - \hat{y})

@@

\section{Experiments}
We experimentally evaluate the approach discussed in Section~\ref{sec:gradient-reverse} 
on a noisy speech recognition dataset Aurora-4~\cite{parihar2002aurora}. This dataset
is based on the Wall Street Journal corpus (WSJ0 is available at LDC). Noise of 
six categories (airport, babble, car, restaurant, street, train) was introduced to the 
clean data. This noise was added and filtering simulating sampling frequency was applyed 
both to clean and noisy data. 
We used mel-filterbank features with their deltas and delta-deltas spliced
over 5 frames on each side making 11 consecutive frames and 1320 input features in total. 
Global mean and variance normalization was applied to these features.
We did not use any speaker normalization of input features. We pretrained 5 layer
speech recognition network with 2048 rectified linear units at every layer using Adam 
optimization algorithm for 5 epochs. After pretraining, we turned on the invariance term.

The first expriment we performed is the gender invariance. We added a 2 layer network 
branching off the last layer of the pretrained speech recognition network and 
trained it until convergence with SGD. Although, the performance remained unchanged,
the t-SNE visualization in Figure~\ref{fig:tsne} demonstrates that the hidden units
are less separated in the projected space.

Next, we trained a model for the noise condition invariance. We used an architecture
similar to one described above with a binary invariance classifier target denoting 
whether the utterance is clean or noisy. We oversampled the noisy utterances to 
address the fact that the labels are not ballanced, the clean subset has approximately 4400
when every noisy condition subset contains only 443 utterances. The results are
summarized in Table~\ref{tab:results}.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.5\linewidth}
        \centering
        \includegraphics[width=\linewidth]{tsne-original.png}
    \end{subfigure}%
    \begin{subfigure}[b]{0.5\linewidth}
        \centering
        \includegraphics[width=\linewidth]{tsne-invariant.png}
    \end{subfigure}%
    \caption{t-SNE visualization of the last hidden layer using the clean development
    subset. On the left, the baseline
    model. On the right, the model trained with the adversarial objective. Colors 
    represent genders.}
    \label{fig:tsne}
\end{figure}

\begin{table}
    \centering
    \caption{Results for word error rate (WER\%) on Aurora-4 dataset.}
    \label{tab:results}
    \begin{tabular}{r|cc|cc}
    \multirow{2}{*}{Subset} & \multicolumn{2}{c}{Orig}  & \multicolumn{2}{|c}{Inv}\\ 
               & wv1  & wv2  & wv1  & wv2 \\
    \hline 
    clean      & 4.04 & 6.76 & 5.03& 7.53 \\
    airport    & 7.44 & 15.36& 8.22& 16.03\\
    babble     & 7.86 & 15.88& 9.06& 17.82\\
    car        & 4.17 & 8.99 & 5.21& 9.81 \\
    restaurant & 9.25 & 17.32& 12.2& 19.60\\
    street     & 7.79 & 16.14& 9.34& 17.56\\
    train      & 8.18 & 17.26& 9.60& 18.66\\
    \hline
    Average    & \multicolumn{2}{c}{10.46}& & 
    \end{tabular}
    %TODO: add numbers!
\end{table}

\section{Conclusion}

@@

We would like to thank Yaroslav Ganin, David Warde-Farley, Samuel Thomas for insightful discussions,
developers of Theano~\cite{2016arXiv160502688short}, Blocks, and Fuel~\cite{MerrienboerBDSW15} 
for great instruments.

>>>>

% Title.
% ------
\title{Invariant Representations for Noisy Speech Recognition}
%
% Single address.
% ---------------
%\name{Dmitriy Serdyuk $^{\dagger}$\thanks{Dmitriy Serdyuk performed the work
%  while at IBM}, Kartik Audhkhasi$^{\star}$, Phil\'emon Brakel$^{\dagger}$, Bhuvana Ramabhadran$^{\star}$, Yoshua Bengio$^{\dagger}$\thanks{Yoshua Bengio is a CIFAR Fellow.}}
%\address{$^{\dagger}$ Universit\'e de Montr\'eal\\
%         $^{\star}$ IBM}
\author{
  Dmitriy Serdyuk\thanks{Dmitriy Serdyuk performed the work 
    during an internship at IBM Watson} \\
  MILA\\
  Universit\'e de Montr\'eal\\
  Montr\'eal, QC H3T 1J4 \\
  \texttt{serdyuk@iro.umontreal.ca} \\
  \And
  Kartik Audhkhasi \\
  IBM Watson Research Center \\
  Yorktown Heights, NY 10598\\
  \texttt{kaudhkha@us.ibm.com} \\
  \And
  Phil\'emon Brakel \\
  MILA\\
  Universit\'e de Montr\'eal\\
  Montr\'eal, QC H3T 1J4 \\
  \texttt{pbpop3@gmail.com} \\
  \And
  Bhuvana Ramabhadran\\
  IBM  Watson Research Center\\
  Yorktown Heights, NY 10598\\
  \texttt{bhuvana@us.ibm.com} \\
  \And
  Samuel Thomas\\
  IBM  Watson Research Center\\
  Yorktown Heights, NY 10598\\
  \texttt{bhuvana@us.ibm.com} \\
  \And
  Yoshua Bengio\\
  MILA, CIFAR Fellow\\
  Universit\'e de Montr\'eal\\
  Montr\'eal, QC H3T 1J4 \\
  \texttt{findme@inter.net} \\
}

@@

@@

    The network is trained with three goals: the hidden representation $h$ should
    be helpful for the target classifier, but harmful for the domain classifier,
    the domain classifier should have a good classification accuracy. More 
    formally, we define the objective as
    \begin{equation}
        L = L(\hat{y}, y; \theta_R, \theta_E) + 
        \alpha L(\hat{d}, d; \theta_D) -
        \beta L(\hat{d}, d; \theta_E),
    \end{equation}
    where $y$ is the ground truth class, $d$ is the domain label, corresponding
    hatted variables are the network predictions, $\theta$ are the network 
    parameters, and $E, R, D$ are the subsets of parameters in the encoder,
    recogniser and the domain adaptation networks respectively. The hyper-parameters
    $\alpha$ and $\beta$ denote the influence of every part of the cost.

\subsection{Invariant Represenations for Speech Recognition}
\label{sec:invariant-speech}

In this work we focus on speech recognition using DNN-HMM hybrid systems. The
recording conditions, speaker identity, or gener might be considred to be domains
and adptation task corresponds to making the hidden layer representations be 
invariant with respect to those conditions. We hypothesize that this method of
training helps to generalize on unseen conditiions and requires only very small
amount of labeling the conditions, see Section~\ref{sec:experiments} for experimental
evaluation.

The model we use is depicted on Figure~\ref{fig:model}. It is a feedforward neural network
which is trained to predict the HMM state tweaked with a branch for the invariance
training. This branch is discarded in the testing phase. In our experiments we
used the noise condition as the domain label merging all noise type to one label
and clean type to another.

The particular cost we used for our training is
    \begin{equation}
        L = L_1(\hat{y}, y; \theta_R, \theta_E) + 
        L_2(\hat{d}, d; \theta_D) +
        \beta L_3(\hat{d}, d; \theta_E),
        \label{eq:cost}
    \end{equation}
where $L_3 = - d\log(1 - \hat{d}) - (1-d)\log(\hat{d})$ maximizes the probablity
of an incorrect domain classification. The terms for $L_1$ and $L_2$ are 
regular cross-entropy terms which are minimized with respect to the parameters
after semicolon.
We don't use any coefficient for first two terms because it is equivalent to 
changing the learning rate.

@@

\section{Experiments}
\label{sec:experiments}
We experimentally evaluate the approach discussed in Section~\ref{sec:gradient-reverse} 
on a noisy speech recognition dataset Aurora-4~\cite{parihar2002aurora}. This dataset
is based on the Wall Street Journal corpus (WSJ0 is available at LDC). It contains noise of 
six categories which was added to clean
data. Every clean and noisy utterance filtered to simulate sampling frequency. The train
set contains approximately 4400 clean utterances and 446 utterances of each noise condition,
in total 2676 noisy utterances.
The test
set contains clean data, the same 6 noises, and data recorded with a different microphone
both clean and noise corrupted.

We used 16kHz training audio. We extracted mel-filterbank features with their deltas and 
delta-deltas spliced over 5 frames on each side making 11 consecutive frames and 1320 input 
features in total. Global mean and variance normalization was applied to these features.
We did not use any speaker normalization of input features. We trained 6 layer
speech recognition network with 2048 rectified linear units at every layer using momentum
accelerated stochastic gradient descent for 15 epochs. We annealed the learning rate accroding 
to the development set dividing it by 2 when the cross entropy did not improve.

In order to evaluate the impact of our method on generalization to new, unseen noises 
we performs 6 groups of experiments with different set of seen noises. We always train
on clean data and add noise conditions one-by-one in order: airport, babble, car, 
restaurant, street, train. The last group includes all noises therefore matches the
standard multi condition training setup. For every group of experiments we trained the
baseline and the invariance model where we branch out at 4th layer to a logistic 
binary classifier clean versus noisy data. Due to the imballance between clean and
noisy parts of the dataset, we had to oversample noisy frames, every minibatch contained
half of clean and half of noisy data points.

The results are summarized in Table~\ref{tab:results}. 
In our experiments we did not use layerwise pretraining not uncommonly used for small
datasets due to the difficulties of combining it with the invariance training. Our preliminary
experiments show that the network is not able to leave the parameter region it reaches after 
pretraining.

%\begin{figure}
%    \centering
%    \begin{subfigure}[b]{0.5\linewidth}
%        \centering
%        \includegraphics[width=\linewidth]{tsne-original.png}
%    \end{subfigure}%
%    \begin{subfigure}[b]{0.5\linewidth}
%        \centering
%        \includegraphics[width=\linewidth]{tsne-invariant.png}
%    \end{subfigure}%
%    \caption{t-SNE visualization of the last hidden layer using the clean development
%    subset. On the left, the baseline
%    model. On the right, the model trained with the adversarial objective. Colors 
%    represent genders.}
%    \label{fig:tsne}
%\end{figure}

\begin{table}
    \centering
    \caption{Average word error rate (WER\%) on Aurora-4 dataset on all test conditions,
        including seen and unseen noise and unseen microphone. First column
        is the number of noise conditions used for the training set. The last row is a 
        preliminary experiment with layer-wise pretraining close to state-of-the-art
        model and a corresponding invariance training starting with a pretrained model.}
    \label{tab:results}
    \begin{tabular}{r|cc}
    %\multirow{2}{*}{Subset} & \multicolumn{2}{c}{Orig}  & \multicolumn{2}{|c}{Inv}\\ 
    %           & wv1  & wv2  & wv1  & wv2 \\
    Noise       &Invariant    &Baseline\\
    \hline
    1           &16.36*       &18.14 \\
    2           &15.56*       &17.39 \\
    3           &14.29        &14.67 \\
    4           &13.61        &13.84 \\         
    5           &13.41        &13.02 \\         
    6           &12.62        &12.60 \\
    \hline\hline
    6, pre-train&11.88        &11.99
    \end{tabular}
    %TODO: add numbers!
\end{table}

\section{Discussion}
    We demonstrate that the invariance training helps to generalize to unseen noise
    conditions and improves in

    The task of speech recognition differs from the image recognition
    adaptation in an aspect of the domain adaptation network training dynamics. 
    While the most common problem for the GAN type of networks trained on images 
    is overfitting, the noise type classification networks suffers from 
    underfitting. This makes the gradient of the $L_3$ term in Eq.~\ref{eq:cost}
    unrealiable and noisy.

@@

We would like to thank Yaroslav Ganin, David Warde-Farley for insightful discussions,
developers of Theano~\cite{2016arXiv160502688short}, Blocks, and Fuel~\cite{MerrienboerBDSW15} 
for great instruments.
